---
title: "Average_gain_differ"
output: html_document
Author: 'Porhay Rouen'
---

# Loading package
```{r}
library(nycflights13)
library(ggplot2)
library(dplyr)
```

Introduction

This project analyzes net gain per flight for United Airlines (UA). Net gain is defined as departure delay minus arrival delay, which shows how much quicker a flight arrived than planned. We will explore whether net gain differs for flights that departed late versus on time, and for flights that departed more than 30 minutes late.

## check few data
```{r}
head(flights)
```

# Filter United Airlines flights and compute net gain
```{r}

ua_flights <- flights %>%
filter(carrier == "UA") %>%
mutate(
net_gain = dep_delay - arr_delay,
departed_late = dep_delay > 0,        # any positive delay
departed_30plus = dep_delay > 30      # more than 30 minutes late
) %>%
filter(!is.na(net_gain))

```

#Exploratory Data Analysis (EDA)

```{r}
# Summary of net gain

summary(ua_flights$net_gain)

# Histogram of net gain

ggplot(ua_flights, aes(net_gain)) +
geom_histogram(bins = 40, fill = "skyblue", color = "black") +
labs(
title = "Distribution of Net Gain for UA Flights",
x = "Net Gain (minutes)",
y = "Count"
)

```

# Compare Late vs Not Late

```{r}
# Compute mean net gain

mean_late <- ua_flights %>%
group_by(departed_late) %>%
summarise(mean_gain = mean(net_gain))

# Bar plot

ggplot(mean_late, aes(x = departed_late, y = mean_gain, fill = departed_late)) +
geom_col(width = 0.5) +
geom_text(aes(label = round(mean_gain,2)), vjust = -0.5) +
scale_fill_manual(values = c("skyblue", "orange")) +
labs(
title = "Average Net Gain: Late vs Not Late",
x = "Departed Late?",
y = "Mean Net Gain (minutes)"
) +
theme_minimal()


```
Flights that departed on time gained 9.27 minutes on average, while late flights gained 7.54 minutes. This difference is statistically significant, as confirmed by the t-test below.

```{r}
# T-test: Late vs Not Late

t_test_late <- t.test(net_gain ~ departed_late, data = ua_flights)
t_test_late

```

# Compare >30 Min Late vs Others

```{r}
# Compute mean net gain

mean_late30 <- ua_flights %>%
group_by(departed_30plus) %>%
summarise(mean_gain = mean(net_gain))

# Bar plot

ggplot(mean_late30, aes(x = departed_30plus, y = mean_gain, fill = departed_30plus)) +
geom_col(width = 0.5) +
geom_text(aes(label = round(mean_gain,2)), vjust = -0.5) +
scale_fill_manual(values = c("skyblue", "orange")) +
labs(
title = "Average Net Gain: >30 Min Late vs Others",
x = "Departed >30 Min Late?",
y = "Mean Net Gain (minutes)"
) +
theme_minimal()

```

Flights delayed over 30 minutes gained 6.86 minutes on average, while other flights gained 8.70 minutes. The difference is statistically significant according to the t-test.

```{r}
# T-test: >30 Min Late vs Others

t_test_late30 <- t.test(net_gain ~ departed_30plus, data = ua_flights)
t_test_late30

```

